On the presentation of orthogonal groups over polynomial rings

### 2019-05-31 dl-paper.tex
Proved R3_{1,0}

\documentclass[oneside, 8pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true,unicode]{hyperref}
\usepackage[hyperref=true, backend=biber, bibencoding=utf8, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}
%\usepackage[notref,notcite]{showkeys}

@@

\begin{df} \label{defP0} Denote by $P_\alpha(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following elements of $\St(\Phi, A[X])$ under the
 map $j_+ \colon \overline{\St}(\Phi, A[X], M[X]) \to \St(\Phi, A[X, X^{-1}])$:
\begin{enumerate} \item $z_{\beta}(Xf, \xi), \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha;$
 \item $x_{-\alpha}(X^2f);$
 \item $x_{\alpha}(f)$. \end{enumerate}
In the above formulae and for the rest of this subsection $f \in M[X]$ and $\xi \in A[X]$. \end{df}

\begin{rem}\label{rem:c-DS} It is easy to check that $P_\alpha(0)$ contains the image of $Z_\alpha(A, M)$ under the natural embedding of $\St(\Phi, A, M) \hookrightarrow \St(\Phi, A[X, X\inv])$.
In particular, it contains the relative Dennis--Stein and Steinberg symbols.
It is also clear that $P_\alpha(0)$ contains the elements $c_{\beta}(f, X\xi)$ for all $\beta \in Z(\alpha)$
 (they can be factored as products of two elements of type 2 or 3). \end{rem}

Denote by $P_\alpha(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_\alpha(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$.
The following lemma shows that the resulting subgroup is sufficiently large.
\begin{lemma} \label{Pstar-large} The subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$ hence by~\cref{thm:Stepanov} $P_\alpha(*)$ contains whole $j_+(\overline{\St}(\Phi, A[X], XM[X]))$. The required assertion now follows from~\cref{Kgen-strong} and the preceding remark. \end{proof}

\begin{rem} \label{Pstar-char} The above lemma allows us to characterize $P_\alpha(*)$ as follows: it consists precisely of images under $j_+$ of all elements 
 $g \in \overline{\St}(\Phi, A[X], M[X])$ for which $ev_{X=0}^*(g)$ lies in the subgroup $Z_\alpha(A, M)$. \end{rem}

\begin{lemma}\label{P0_normal} The subgroup $P_\alpha(0)$ is normal in $P_\alpha(*)$. In particular, there is an exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar[r]^{p_\alpha} & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type 3 and 4.

@@

\begin{rem} \label{rem:palpha} From the two preceding lemmas it follows that the value of the function $p_\alpha$ from the statement of~\cref{P0_normal} on an element $g\in K(A[X], M[X])$ can be computed via the following procedure. 
Start with any presentation of $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta\in Z_0(\alpha)$. 
Now pick among these factors those that correspond to the root $\beta = -\alpha$ (i.\,e. pick all factors $z_{-\alpha}(Xf_i, \xi_i)$). 
Now $p_\alpha(g)$ is precisely the sum of constant terms of the polynomials $f_i$. \end{rem}

For $m \in M$ denote by $P_\alpha(m)$ the coset $P_\alpha(0) \cdot x_{-\alpha}(mX)$.
From the preceding lemma it follows that $P_\alpha(*)$ coincides with the union of all $P_\alpha(m)$, $m\in M$. This allows us to make the following definition.

@@

\begin{lemma}
 The operators $T_\alpha$ satisfy $\text{R3}_{1,0}$.
\end{lemma}
\begin{proof}
 We have to show that $[T_\alpha(aX^{-1}), T_\beta(b)](g, h, u) = (g, h, u)$ for $\beta\in Z(\alpha)$, $a, b\in A$.
 \begin{multline}
  T_\beta(b) T_\alpha(aX^{-1}) T_\beta(-b) T_\alpha(-aX^{-1})(g, h, u) = \\
  = T_\beta(b) T_\alpha(aX^{-1}) \left(x_\beta(-b) x_\alpha(-aX^{-1}) \cdot g, h \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1-am}\right) \cdot x_\beta(b), u(1-am)\right)
 \end{multline}
\end{proof}

>>>>

\documentclass[oneside, 8pt]{amsart}
\usepackage{amscd, amsmath, amssymb, amsthm, amsfonts, amstext, verbatim, mathtools, xfrac, microtype, nameref, thmtools}
\usepackage[breaklinks=true,unicode]{hyperref}
\usepackage[hyperref=true, backend=biber, bibencoding=utf8, citestyle=numeric-comp, sortlocale=en_US, url=false, doi=false, eprint=true, maxbibnames=4]{biblatex}
\usepackage[capitalize]{cleveref}
\usepackage[matrix,arrow,curve]{xy}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage[notref,notcite]{showkeys}

@@

\begin{df} \label{defP0}
Let $M$ be an ideal of a local ring $A$. We denote by $P_{\alpha, M}(0)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by the images of the following elements of $\St(\Phi, A[X])$ under the
 map $j_+ \colon \overline{\St}(\Phi, A[X], M[X]) \to \St(\Phi, A[X, X^{-1}])$:
\begin{enumerate} \item $z_{\beta}(Xf, \xi), \beta \in \Phi \text{ such that }\alpha + \beta \in \Phi\text{ or } \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, X\xi),\ \alpha - \beta \in \Phi;$
 \item $z_{\beta}(f, \xi),\ \beta \perp \alpha;$
 \item $x_{-\alpha}(X^2f);$
 \item $x_{\alpha}(f)$. \end{enumerate}
In the above formulae and for the rest of this subsection $f \in M[X]$ and $\xi \in A[X]$.

We also denote by $P_{\alpha, M}(*)$ the subgroup of $\St(\Phi, A[X, X^{-1}])$ generated by $P_{\alpha, M}(0)$ and the elements $x_{-\alpha}(mX)$, $m \in M$.
\end{df}

Typically, the ideal $M$ from the above definition is precisely the maximal ideal of $A$, however, sometimes (e.\,g. in~\cref{P0-conj-invariant}) we also use the subgroups defined above for $M=A$.

\begin{lemma}\label{P0_normal} The subgroup $P_{\alpha, M}(0)$ is normal in $P_{\alpha, M}(*)$. In particular, there is a short exact sequence, which is split by the map $m \mapsto x_{-\alpha}(mX)$:
\[\xymatrix{1 \ar[r] & P_\alpha(0) \ar[r] & P_\alpha(*) \ar@{->>}[r]^{p_\alpha} & M^+ \ar@<5pt>@{-->}[l] \ar[r] & 1}.\] \end{lemma}
\begin{proof} We need to verify that the conjugate by $x_{-\alpha}(mX)$ to every generator $g$ of $P_\alpha(0)$ listed in~Definition~\ref{defP0} belongs to $P_\alpha(0)$.
The assertion is obvious for the generators of type 3 and 4.

@@

\begin{rem}\label{rem:c-DS} It is easy to check that $P_{\alpha, M}(0)$ contains the image of $Z_\alpha(A, M)$ under the natural embedding of $\St(\Phi, A, M) \hookrightarrow \St(\Phi, A[X, X\inv])$.
In particular, it contains the relative Dennis--Stein and Steinberg symbols.
Notice that $P_\alpha(0)$ also contains the elements $c_{\beta}(f, X\xi)$ for all $\beta \in Z(\alpha)$ (they can be factored as products of two elements of type 2 or 3). \end{rem}

The following lemma shows that $P_{\alpha, M}(*)$ is sufficiently large.
\begin{lemma} \label{Pstar-large} Suppose that $(A, M)$ is a local pair. Then the subgroup $P_\alpha(*)$ contains the image of $K(A[X], M[X])$ in $\St(\Phi, A[X, X^{-1}])$. \end{lemma}
\begin{proof} Clearly, $P_\alpha(*)$ contains the elements $x_\beta(Xf)$ for all $\beta \in \Phi$ and $z_\beta(Xf, \xi)$ for $\beta \in \Phi \setminus \{\pm \alpha\}$, $f \in M[X]$, $\xi\in A[X]$ hence by~\cref{thm:Stepanov} $P_\alpha(*)$ contains whole $j_+(\overline{\St}(\Phi, A[X], XM[X]))$. The required assertion now follows from~\cref{Kgen-strong} and the preceding remark. \end{proof}

\begin{rem} \label{Pstar-char} The above lemma allows us to characterize $P_{\alpha, M}(*)$ as follows: it consists precisely of images under $j_+$ of the elements 
 $g \in \overline{\St}(\Phi, A[X], M[X])$ for which $ev_{X=0}^*(g)$ lies in the subgroup $Z_\alpha(A, M)$. \end{rem}

\begin{rem} \label{rem:palpha} It follows from~\cref{P0_normal} and~\cref{Pstar-large} that the value of the function $p_\alpha$ from the statement of~\cref{P0_normal} on an element $g\in K(A[X], M[X])$ can be computed via the following procedure. 
Start with any presentation of $g$ as a product of elements $z_{\beta}(Xf, \xi)$ for $\beta \in \Phi$ and $c_\delta(f, X\xi)$ for some fixed $\delta\in Z_0(\alpha)$. 
Now pick among these factors those that correspond to the root $\beta = -\alpha$ (i.\,e. pick all factors $z_{-\alpha}(Xf_i, \xi_i)$). 
Now $p_\alpha(g)$ is precisely the sum of constant terms of the polynomials $f_i$. \end{rem}

\begin{lemma} \label{P0-conj-invariant} Suppose $(A, M)$ is a local pair. Then for any $\beta \in Z(\alpha)$ and $b \in A$ the subgroups $P_{\alpha, M}(0)$ and $P_{\alpha, M}(*)$ are stable under conjugation by $x_\beta(b)$. \end{lemma}
\begin{proof}
Notice that both $Z_\alpha(A, M)$ and $K(A[X], M[X])$ are stable under the specified conjugation, which implies the assertion for $P_{\alpha, M}(*)$. 
To obtain the assertion for $P_{\alpha, M}(0)$ consider the following commutative diagram. \[ \xymatrix{ P_{\alpha, M}(0) \ar@{^{(}->}[r] \ar@{^{(}->}[d] & P_{\alpha, M}(*) \ar@{->>}[r]^{p_{\alpha, M}} \ar@{^{(}->}[d] & M^+ \ar@{^{(}->}[d] \\ P_{\alpha, A}(0) \ar@{^{(}->}[r] & P_{\alpha, A}(*) \ar@{->>}[r]^{p_{\alpha, A}} & A^+} \]
Notice that $x_\beta(b) \in P_{\alpha, A}(0)$ therefore for $g \in P_{\alpha, M}(0)$ one has \[p_{\alpha, M}(x_\beta(b) \cdot g \cdot x_\beta(-b)) = p_{\alpha, A}(x_\beta(b) \cdot g \cdot x_\beta(-b)) = p_{\alpha, A}(g) = 0,\] which implies the assertion.\end{proof}

For the rest of this section $M$ denotes the maximal ideal of $A$, so everywhere below we abridge $P_{\alpha, M}(*)$ (resp. $P_{\alpha, M}(0)$) to just $P_{\alpha}(*)$ (resp. $P_{\alpha}(0)$). For $m \in M$ denote by $P_\alpha(m)$ the coset $P_\alpha(0) \cdot x_{-\alpha}(mX)$.
From~\cref{P0_normal} it follows that $P_\alpha(*)$ coincides with the union of all $P_\alpha(m)$, $m\in M$. This allows us to make the following definition.

@@

\begin{lemma}
 The operators $T_\alpha$ satisfy $\text{R3}_{1,0}$.
\end{lemma}
\begin{proof}
We have to show that $[T_\alpha(aX^{-1}), T_\beta(b)](g, h, u) = (g, h, u)$ for $\beta\in Z(\alpha)$, $a, b\in A$. Suppose $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$. It follows from~\cref{P0-conj-invariant} that $P_\alpha\left(\tfrac{m}{1-am}\right)$ is stable under conjugation with $x_\beta(b)$, therefore from~\cref{SR:additivity} we obtain that 
\begin{gather*}
\begin{multlined} \nonumber
  T_\beta(b) T_\alpha(aX^{-1}) T_\beta(-b) T_\alpha(-aX^{-1})(g, h, u) \sim \\
  \sim T_\beta(b) T_\alpha(aX^{-1}) \left(x_\beta(-b) x_\alpha(-aX^{-1}) \cdot g, h \cdot x_{\alpha}\left(\tfrac{aX^{-1}}{1-am}\right) \cdot h^{-1}_\alpha(1-am) \cdot р x_\beta(b), u(1-am)\right) = \\ =
  \left( [x_\beta(b), x_{\alpha}(aX^{-1}] \cdot g, h \cdot h', u] \right) \sim \left( g, h, u\right), \end{multlined} \\
\begin{multlined} \nonumber \text{ where }
 h' = x_{\alpha}\left(\tfrac{aX^{-1}}{1-am}\right) \cdot h^{-1}_\alpha(1-am) \cdot р x_\beta(b) \cdot x_{\alpha}(-a(1-am)X^{-1}) \cdot x_{\beta}(-b) = \\ = h_\alpha^{-1}(1-am) \left[x_\alpha(\tfrac{aX^{-1}}{1-am}), x_\beta(b)\right] = h_\alpha^{-1}(1-am). \qedhere \end{multlined} \end{gather*}
